---
title: 'Metalinguistic Analysis'
author: "Mark Newman"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, echo = F, results = 'hide', error = F, warning = F, message = F}
rm(list = ls())
options(scipen = 999)

libraries <- c('dplyr', 'ggpubr')
lapply(libraries, library, character.only = TRUE)
```

A Metalinguistic Analysis compares the lingustic features of a corpus accross some metadata segment.
The features are extracted on a document by document basis.
Then a MANOVA is preformed comparing the segments.

In order to perform a Metalinguistic Analysis, documents must first be obtained.
This can be done by any means.
Following data collection, the documents need to be processed and the features extracted.
R is _OK_ enough at this task, but as the document count increases, R _really_ slows down.
Because of this, the feature extraction step is preformed in Pure Python.
The walkthrough of this can be found in `../code/01-MetalinguisticAnalysisPipeline.md`

# Data format
 
For this recipe, 3 files are expected:

* A `metadata.csv` file containing metadata data about the individual documents.
  This file will need to include at least 2 columns: (1) the document's id, and (2) the segment.
* A `cadence.csv` file containing frequency count measures.
  This file will contain 4 columns
* A `diversity.csv` file containing metadata data about the individual documents.



## Question

```{r, results = 'hold'}
analysis = 'Metalinguistic'
```
